---
title: "SNP clinality"
date: "`r date()`"
---

```{r setup, include=FALSE}
library(pander)
library(raster)
library(maptools)
# library(phonTools)  # for fastacf
options(scipen=3)
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
# locations and labels
tortdir <- gsub("tortoisescape.*","tortoisescape",getwd())
coord.obj <- load(file.path(tortdir,"tort_272_info","geog_coords.RData"))
coords <- get(coord.obj)
tort.ids <- row.names(coords)
nindivs <- length(tort.ids)
```

```{r snp_info, include=FALSE, cache=TRUE}
datadir <- "."
countfile <- file.path(datadir,"272torts_snp1e6_minmapq20minq30_map2kenro.counts.gz")
bincountfile <- "272torts_snp1e6_minmapq20minq30_map2kenro.counts.bin"
# scaffold info
scaf.info <- read.table(file.path(datadir,"GopAga1.0.softmasked.scafflength"),header=TRUE,stringsAsFactors=FALSE)
names(scaf.info) <- c("chr","length")
# maf <- read.table(file.path(datadir,"272torts_snp1e6_minmapq20minq30_map2kenro.mafs.gz"),header=TRUE)
pos <- read.table(file.path(datadir,"272torts_snp1e6_minmapq20minq30_map2kenro.pos.gz"),header=TRUE,stringsAsFactors=FALSE)
scaf.info <- subset(scaf.info,chr%in%pos$chr)
pos$chr <- factor(pos$chr,levels=scaf.info$chr)
# scaffolds
scaf.counts <- table( pos$chr )
minscaf <- 400  # minimum number of snps per scaffold
long.scafs <- names(scaf.counts)[scaf.counts>=minscaf]
# apply filters
mindepth <- 150 # minimum totalDepth
maxdepth <- 500 # maximum totalDepth
goodones <- ( ( pos$totDepth >= mindepth ) & ( pos$totDepth <= maxdepth ) )
# minind <- 250   # minimum nInd -- RESULTS IN NO SNPs
# goodones <- goodones & ( maf$nInd >= minind )
goodones <- goodones & ( pos$chr %in% long.scafs )
```

```{r plot_scafs, include=FALSE, cache=TRUE}
# whatnot to plot things 'along the genome'
scaf.len <- as.numeric(scaf.info$length[match(levels(pos$chr),scaf.info$chr)])
names(scaf.len) <- levels(pos$chr)
scaf.breaks <- c(0,cumsum(as.numeric(scaf.len)))
scaf.pos <- scaf.breaks[-1]-diff(scaf.breaks)/2
scafplot <- function (y, scafs=TRUE,
        pch=20, cex=0.5,
        col=adjustcolor("black",0.75),
        xlab='cumulative position (Mb)',
        ylab='', add=FALSE, ...) {
    # y is a vector of length nrow(pos)
    # whose non-NA entries will be plotted
    the.scafs <- levels(pos$chr)[scafs]
    scaf.breaks <- c(0,cumsum(as.numeric(scaf.len[scafs])))
    use.these <- is.finite(y)
    xx <- scaf.breaks[match(pos$chr[use.these],the.scafs)] + pos$pos[use.these]
    yy <- y[use.these]
    if (!add) plot( xx/1e6, yy, type='n', 
                xlab=xlab, ylab=ylab, ... )
    rect(xleft=scaf.breaks[-length(scaf.breaks)]/1e6, xright=scaf.breaks[-1]/1e6,
            ybottom=par('usr')[3], ytop=par("usr")[4], col=c(adjustcolor('blue',0.2),NA),
            border=NA )
    points( xx/1e6, yy, pch=pch, cex=cex, col=col )
}
scafpoints <- function (...) { scafplot(...,add=TRUE) }
```

```{r count_connection, include=FALSE}
# the counts themselves
bincount <- file(bincountfile,open="rb")
attr(bincount,"nindivs") <- nindivs
attr(bincount,"nbytes") <- 1
read_bincounts <- function (bincount,lines) {
    line.coords <- 4*(lines-1)*attr(bincount,"nbytes")*attr(bincount,"nindivs")
    output <- matrix( integer(4*length(lines)*attr(bincount,"nindivs")), nrow=length(lines) )
    for (k in seq_along(lines)) {
        seek(bincount,line.coords[k])
        output[k,] <- readBin( bincount, what=integer(),
                              n=4*attr(bincount,"nindivs"),
                              size=attr(bincount,"nbytes"),
                              signed=(attr(bincount,"nbytes")>2) )
    }
    output[output>=256^attr(bincount,"nbytes")-1] <- NA
    return(output)
}
```

# Data

We are only looking at scaffolds with at least `r minscaf` SNPs per scaffold,
and SNPs with total depth between `r mindepth` and `r maxdepth`;
this is a proportion `r mean(goodones)` of the SNPs.
Overall, we've got a depth of about `r mean(pos$totDepth[pos$totDepth<maxdepth])`;
on the longer scaffolds this is `r mean(pos$totDepth[pos$totDepth<maxdepth & pos$chr%in%long.scafs])`;


# Computation

We want to compute the correlation and covariance of each site with a vector, $v$ (say, PC1).
If at a given site:

- $c(i) =$ major allele count in indiv $i$
- $n(i) =$ sample size, i.e. coverage of indiv $i$
- $p(i) =$ $c(i)/n(i)$
- $v(i) =$ vector of weights

then we need these things:

- $A = \sum_i p(i)$
- $B = \sum_i p(i) v(i)$
- $C = \sum_{i : p(i)>0} v(i)$
- $D = \sum_{i : p(i)>0} v(i)^2$
- $N = \#\{ i : n(i) > 0 \}$

Then, we estimate the covariance by
$$  ( B - A C / N ) / N $$
and the correlation by
$$
  ( \text{covariance} ) / \sqrt{ ( ( A - A^2 / N ) / (N-1) ) ( ( D - C^2 / N ) / (N-1) ) }
  = ( B N - A C ) (N-1) / N \sqrt{ ( A N - A^2 ) ( D N - C^2 ) }
$$

The five columns of the pcXcounts files are $A$, $B$, $C$, $D$, and $N$, respectively.

Note that from these data we can also look for clinality in the **missing** data.
Suppose that

- $z(i) = 1$ if the allele is missing in individual $i$ and $=0$ otherwise;
- $C = \sum_i v(i) z(i)$
- $N = \sum_i z(i)$
- $n_I =$ the number of individuals
- $\bar v = \sum_i v(i) / n_I$
- $\sigma^2_v = \sum_i (v(i)-\bar v)^2 / (n_I-1)$

and so the covariance of $z$ with $v$ is
$$  ( C - N \sum_i v(i) / n_I ) / n_I $$
and the correlation is
$$  ( \text{covariance} ) / \sqrt{ ( N / n_I )( 1 - N/n_I ) \sigma^2_v } . $$

```{r pc_count_connection, include=FALSE, cache=TRUE}
pcs <- read.csv(file.path(tortdir,"tort_272_info","pcs.csv"),header=TRUE,stringsAsFactors=FALSE)
get_pc_cors <- function (pc.num){
    # correlation of PCs with counts
    countsuffix <- sprintf(".pc%dcounts.5bin",pc.num)  # .5bin means binary, five columns
    pccountfile <- gsub(".counts.gz",countsuffix,countfile)
    pccountheaderfile <- paste0(pccountfile,".header")
    pc.header <- scan(pccountheaderfile,what='char')
    pc.con <- pipe(paste("cat",pccountfile), open="rb")
    on.exit(close(pc.con),add=TRUE)
    read_chunk <- function (blocksize) {
        # do this multiple times to read in chunks
        pc.counts <- readBin(pc.con,what="numeric",n=length(pc.header)*blocksize)
        dim(pc.counts) <- c(length(pc.header),length(pc.counts)/length(pc.header))
        pc.counts <- t(pc.counts)
        colnames(pc.counts) <- pc.header
        return(pc.counts)
    }
    umat <- read_chunk(nrow(pos))
    ## do correlations
    # A is : sum_freq  
    # B is : freq_prod 
    # C is : sum_weights
    # D is : sum_weights_sq
    # N is : num_nonzero 
    pcvec <- pcs[[paste0("PC",pc.num)]]
    pcvec.mean <- umat[,"sum_weights"] / umat[,"num_nonzero"]
    pcvec.var <- ( umat[,"sum_weights_sq"] - pcvec.mean * umat[,"sum_weights"] ) / (umat[,"num_nonzero"]-1) 
    snp.freq <- umat[,"sum_freq"] / umat[,"num_nonzero"]
    # snp.var <- snp.freq * ( 1 - snp.freq )
    pc.cov <- ( umat[,"freq_prod"] / umat[,"num_nonzero"] - pcvec.mean * snp.freq )
    pc.cor <- pc.cov / sqrt( snp.freq * ( 1 - snp.freq ) * pcvec.var )
    ## and, for NAs
    total.pc.mean <- mean(pcvec)
    total.pc.var <- var(pcvec)
    na.freq <- umat[,"num_nonzero"] / nindivs
    na.cov <- ( umat[,"sum_weights"] / nindivs - total.pc.mean * na.freq )
    na.cor <- na.cov / sqrt( na.freq * ( 1 - na.freq ) * total.pc.var )
    # don't need umat any more
    rm(umat); gc()
    return( data.frame(
                snp.freq=snp.freq,
                snp.cor=pc.cor,
                na.freq=na.freq,
                na.cor=na.cor
           ) )
}
pc1.df <- get_pc_cors(1)
pc2.df <- get_pc_cors(2)
```

# Summary stats

Here is the histogram of minor allele frequencies,
with and without the filters above:
(recall that these were already selected by angsd as likely polymorphic)
```{r afs}
hist(1-pc1.df$snp.freq,breaks=1000, xlim=c(0,1), main="empirical minor allele frequency")
hist(1-pc1.df$snp.freq[goodones[seq_along(pc1.df$snp.freq)]],breaks=1000, xlim=c(0,1), main="same, on filtered sites")
```

Some more summary statistics.
We don't have the scaffold lengths handy, so we'll just use the maximal position of a SNP:
Here is the density of segregating sites,
mean divergence at segregating sites (calculated as $2p(1-p)$, where $p$ is the minor allele frequency),
and mean divergence overall,
on scaffolds of length at least 10Kb:
```{r sum_stats_by_scaf}
dens.segsites <- scaf.counts/scaf.len
mean.segsites.pi <- tapply( pc1.df$snp.freq, pos$chr, function (x) { mean(2 * x * (1-x)) } )
mean.pi <- tapply( pc1.df$snp.freq, pos$chr, function (x) { sum(2 * x * (1-x)) } ) / scaf.len
usethese <- (scaf.len>10000)
scaf.summaries <- data.frame( "scaffold length"=scaf.len[usethese],
                  "seg. sites per bp"=as.vector(dens.segsites[usethese]),
                  "divergence at seg sites"=mean.segsites.pi[usethese],
                  "divergence"=mean.pi[usethese]
              )
cor(scaf.summaries)
layout(t(1:2))
for (k in 1:ncol(scaf.summaries)) {
    hist(scaf.summaries[[k]], main=names(scaf.summaries)[k], breaks=100)
}
pairs(scaf.summaries)
```

# Correlation coefficients

This plot shows the number of SNPs having at least a certain absolute correlation with PC1
that pass the coverage conditions above:
```{r cor_cdf}
filtered <- ( ( pos$totDepth >= mindepth ) & ( pos$totDepth <= maxdepth ) )
plot( head(sort(abs(pc1.df$snp.cor[filtered]),decr=TRUE),10000), 
        ylab="absolute correlation with PC", ylim=c(0.4,1),
        xlab="SNP order", pch=20, cex=0.5 )
points( head(sort(abs(pc2.df$snp.cor[filtered]),decr=TRUE),10000), 
        col='red', pch=20, cex=0.5 )
legend("topright",pch=20,col=c("black","red"),legend=c("PC1","PC2"))
```

And, here are the distributions of correlation coefficients with the PCs:
```{r cor_hists}
layout(t(1:2))
hist(pc1.df$snp.cor,breaks=1000,xlim=c(-1,1), main="SNP correlations with PC1")
hist(pc2.df$snp.cor,breaks=1000,xlim=c(-1,1), main="SNP correlations with PC2")
hist(pc1.df$snp.cor[goodones[seq_along(pc1.df$snp.cor)]],breaks=1000,xlim=c(-1,1), main="SNP correlations with PC1, filtered")
hist(pc2.df$snp.cor[goodones[seq_along(pc1.df$snp.cor)]],breaks=1000,xlim=c(-1,1), main="SNP correlations with PC2, filtered")
```

Here is correlation against minor allele frequency:
```{r cor_vs_freq, fig.width=2*fig.dim, cache=TRUE, depends=c("snp_info","pc_count_connection")}
layout(t(1:2))
plot( pc1.df$snp.freq[goodones], pc1.df$snp.cor[goodones], pch='.', col=adjustcolor("black",0.5), main="SNPs",
        xlim=c(0,1), ylim=c(-1,1), xlab="frequency", ylab="correlation with PC1" )
plot( pc2.df$snp.freq[goodones], pc2.df$snp.cor[goodones], pch='.', col=adjustcolor("black",0.5), main="SNPs",
        xlim=c(0,1), ylim=c(-1,1), xlab="frequency", ylab="correlation with PC2" )
```

And, frequencies of missingness, 
with the distribution expected if coverage is Poisson(1.2):
```{r na_afs}
xh <- hist(1-pc1.df$na.freq,breaks=1000, xlim=c(0,1), main="Frequency of missing data")
lines((0:nindivs)/nindivs, length(pc1.df$na.freq)*dbinom((0:nindivs),nindivs,exp(-1.2)),col='red',lwd=2)
xh <- hist(1-pc1.df$na.freq[goodones[seq_along(pc1.df$na.freq)]],breaks=1000, xlim=c(0,1), main="same, on filtered sites")
lines((0:nindivs)/nindivs, length(pc1.df$na.freq)*dbinom((0:nindivs),nindivs,exp(-1.2)),col='red',lwd=2)
```
and, distributions of correlation coefficients:
```{r cor_hists_na}
layout(t(1:2))
hist(pc1.df$na.cor,breaks=1000,xlim=c(-1,1), main="NA correlations with PC1")
hist(pc2.df$na.cor,breaks=1000,xlim=c(-1,1), main="NA correlations with PC2")
hist(pc1.df$na.cor[goodones[seq_along(pc1.df$na.cor)]],breaks=1000,xlim=c(-1,1), main="NA correlations with PC1, filtered")
hist(pc2.df$na.cor[goodones[seq_along(pc2.df$na.cor)]],breaks=1000,xlim=c(-1,1), main="NA correlations with PC2, filtered")
```
and, correlation against minor allele frequency:
```{r na_cor_vs_freq, fig.width=2*fig.dim, cache=TRUE, depends="pc_count_connection"}
plot( pc1.df$snp.freq, pc1.df$na.cor, pch='.', col=adjustcolor("black",0.5), main="missingness",
        xlim=c(0,1), ylim=c(-1,1), xlab="frequency", ylab="correlation with PC1" )
plot( pc2.df$snp.freq, pc2.df$na.cor, pch='.', col=adjustcolor("black",0.5), main="missingness",
        xlim=c(0,1), ylim=c(-1,1), xlab="frequency", ylab="correlation with PC2" )
```

# "Along the genome"

Here is a plot of how the larger correlations fall out on the scaffolds:
```{r along_the_genome, fig.width=4*fig.dim}

zz <- ifelse(goodones & abs(pc1.df$snp.cor)>0.7, abs(pc1.df$snp.cor), NA)
scafplot( zz, ylab='abs(correlation)', main='correlation with PC1')
abline(h=0.8,col='red')
# points( scaf.pos/1e6, tapply(zz,pos$chr,mean,na.rm=TRUE), col='orange', cex=0.25 )
# points( scaf.pos/1e6, tapply(ifelse(goodones,abs(pc1.df$snp.cor),NA),pos$chr,mean,na.rm=TRUE), col='red', pch=20, cex=0.25 )

zz <- ifelse(goodones & abs(pc2.df$snp.cor)>0.3, abs(pc2.df$snp.cor), NA)
scafplot( zz, ylab='abs(correlation)', main='correlation with PC2')
# points( scaf.pos/1e6, tapply(zz,pos$chr,mean,na.rm=TRUE), col='orange', cex=0.25 )
# points( scaf.pos/1e6, tapply(ifelse(goodones,abs(pc2.df$snp.cor),NA),pos$chr,mean,na.rm=TRUE), col='red', pch=20, cex=0.25 )
```

Do outliers in correlation with the two PCs coincide?
These scaffolds have at least one outlier in both senses:
(note these cutoffs are above the 99.99% tail)
```{r highcor_table}
mean(pc1.df$snp.cor[goodones]<0.75)
mean(pc2.df$snp.cor[goodones]<0.5)
levels(pos$chr)[which(pmin(tapply( pc1.df$snp.cor>0.75 & goodones, pos$chr, sum ), tapply( pc2.df$snp.cor>0.5 & goodones, pos$chr, sum ) )>0)]
```


# By scaffold

The mitochondrial haplotype has correlation above 0.8 with PC1.
There are `r sum(tapply( pc1.df$snp.cor>0.8 & goodones, pos$chr, sum ))` SNPs
on `r sum(tapply( pc1.df$snp.cor>0.8 & goodones, pos$chr, sum )>0)` scaffolds that pass this threshold.

First, let's look at the top ten longest scaffolds.
**Above zero** are *absolute* correlations with **PC1**, and below  are with **PC2**:
```{r long_scafs, cache=TRUE, depends=c("snp_info","pc_count_connection")}
superlong <- names(sort(scaf.counts,decreasing=TRUE)[1:10])
for (scaf in superlong) {
    plot( pos$pos[pos$chr==scaf & goodones], abs(pc1.df$snp.cor[pos$chr==scaf & goodones]),
            xlab='position', ylab='|correlation|', main=paste('scaffold',scaf),
            cex=0.25, pch=20, col=adjustcolor('black',0.25), ylim=c(-1,1) )
    mtext(c("PC1","PC2"),side=2,at=c(1,-1)*.75)
    points( pos$pos[pos$chr==scaf], -abs(pc2.df$snp.cor[pos$chr==scaf]),
            cex=0.25, pch=20, col=adjustcolor('black',0.25), ylim=c(-1,1) )
    abline(h=0)
}
```


OK, let's see how correlations partition by scaffold.
```{r scaf_stats}
by.scaf <- data.frame(
        snp.1 = tapply( abs(pc1.df$snp.cor[goodones]), pos$chr[goodones], mean ),
        na.1 = tapply( abs(pc1.df$na.cor[goodones]), pos$chr[goodones], mean ),
        snp.2 = tapply( abs(pc2.df$snp.cor[goodones]), pos$chr[goodones], mean ),
        na.2 = tapply( abs(pc2.df$na.cor[goodones]), pos$chr[goodones], mean )
    )
hist( by.scaf$snp.1, xlim=range(0,by.scaf,finite=TRUE), breaks=50, 
        col=adjustcolor("blue",0.5), freq=TRUE, main='mean absolute correlation by scaffold' )
hist( by.scaf$na.1, breaks=50, add=TRUE, freq=TRUE, col=adjustcolor("red",0.5) )
hist( by.scaf$snp.2, breaks=50, add=TRUE, freq=TRUE, col=adjustcolor("purple",0.5) )
hist( by.scaf$na.2, breaks=50, add=TRUE, freq=TRUE, col=adjustcolor("grey",0.5) )
legend("topright", fill=c('blue','red','purple','grey'), legend=outer(c('SNP', 'missingness'),c("PC1","PC2"),paste))
```

## Is there an enrichment of high-correlation SNPs by scaffold?

Now we'll look at numbers of SNPs with correlations with PC1 above certain thresholds,
by scaffold:
```{r high_cor_by_scaf, fig.height=2*fig.dim}
cor.breaks <- seq(.75,.9,by=.05)
num.highcors <- sapply( cor.breaks, function (cb) {
            tapply( (pc1.df$snp.cor >= cb) & (pos$totDepth > mindepth) & (pos$totDepth < maxdepth), pos$chr, sum, na.rm=TRUE )
        } )
dens.highcors <- sweep(num.highcors,1,scaf.len,"/")
outliers <- rownames(num.highcors)[num.highcors[,2]>8]
outliers.ind <- match(outliers,levels(pos$chr))
layout(matrix(1:4,nrow=2))
for (k in seq_along(cor.breaks)) {
    plot( scaf.len[scaf.len>4e3], num.highcors[scaf.len>4e3,k], log='x', 
         xlab='scaffold length', ylab='number of snps', main=paste('correlation above', cor.breaks[k]) )
    for (sl in c(.01,.001,.0001)) { abline(0,sl,untf=TRUE)}
    text( scaf.len[outliers.ind], num.highcors[outliers.ind,k], labels=outliers, pos=4 )
}
```
Plots give number of high-correlation SNPs against length;
contour lines in the plots are at constant density.
Here's the plots for the possible outlier scaffolds:
```{r outlier_scafs, fig.height=1.5*fig.dim}
get_scaf_2 <- function (scaf,pc.num) {
    cbind( pos[pos$totDepth>mindepth & pos$totDepth<maxdepth & pos$chr==scaf,2:3],
        get(paste0("pc",pc.num,".df"))[pos$totDepth>mindepth & pos$totDepth<maxdepth & pos$chr==scaf,] )
}
for (this.scaf in outliers) {
    this.pc1 <- get_scaf_2(this.scaf,1)
    this.pc2 <- get_scaf_2(this.scaf,2)
    cor( cbind( abs(this.pc1[,2:6]), abs(this.pc2[,c(4,6)]) ) )
    layout(1:2)
    opar <- par(mar=c(par("mar"),0.1)[c(5,2,3,4)])
    with( this.pc1, plot( pos, abs(snp.cor), pch=20, cex=0.5, xlab='', xaxt='n', ylim=c(-1,1),
                          ylab='correlation', main=paste('SNP correlation,',this.scaf) ) )
    with( this.pc2, points( pos, -abs(snp.cor), pch=20, cex=0.5 ) )
    abline(h=0)
    par(mar=c(opar$mar,0.1)[c(1,2,5,4)])
    with( this.pc1, plot( pos, snp.freq, pch=20, cex=0.5, xlab='position', ylab='major allele frequency') )
    par(opar)
}
```



# Maps of random segments of alleles

```{r plot_allele_setup, include=FALSE}
bases <- c("A","C","G","T")
base.cols <- adjustcolor(1:4,.5)
layer <- raster(file.path(tortdir,"visualization/dem_30"))
player <- function (...) { plot(layer,legend=FALSE,xlab="",ylab="",xaxt="n",yaxt="n",legend.mar=0,box=FALSE,...) }
pc.cols <- adjustcolor( ifelse( pcs$PC1[match(tort.ids,pcs$etort)] > 0, "blue", "purple" ), .25 )
map_alleles <- function (lines,...) {
    counts <- read_bincounts(bincount,lines)
    freqs <- do.call( cbind, tapply( 1:ncol(counts), 0:(ncol(counts)-1)%%4, function (k) { rowSums(sweep(counts,1,rowSums(counts),"/")[,k]) } ) )
    coverage <- do.call( cbind, tapply( 1:ncol(counts), 0:(ncol(counts)-1)%/%4, function (k) { rowSums(counts[,k]) } ) )
    these.pc1.cors <- pc1.df$snp.cor[lines]
    these.pc2.cors <- pc2.df$snp.cor[lines]
    site.info <- pos[lines,]
    for (site in 1:nrow(counts)) {
        player( main=paste(site.info$chr[site],site.info$pos[site]), ... )
        mtext(side=3,line=0.5,
              text=sprintf("depth: %d, freq: %0.3f, nInd: %d, pc1cor: %0.3f, pc2cor: %0.3f",
                 rowSums(counts)[site], # depth
                 max(freqs[site,]),   # frequency
                 rowSums(coverage>0)[site], # num indivs covered
                 these.pc1.cors[site],  # cor with PC1
                 these.pc2.cors[site])  # cor with PC2
              )
        for (k in seq_along(bases)) {
            nn <- k+4*(0:(nindivs-1))
            points(coords,pch=20,cex=counts[site,nn],col=base.cols[k])
        }
        legend('topleft',legend=bases,col=base.cols,pch=20)
    }
}
```

We'll pull out and make maps of contiguous segments of one hundred SNPs for the top few scaffolds, 
as well as the outlying ones;
here are links to those slideshows:
```{r map_scafs, include=FALSE, cache=TRUE}
for (scaf in superlong) {
    scafdir <- paste0(scaf,"_maps")
    dir.create(scafdir,showWarnings=FALSE)
    thisdir <- file.path( scafdir, "centered" )
    dir.create(thisdir,showWarnings=FALSE)
    scaf.lines <- which( pos$chr==scaf )
    do.lines <- scaf.lines[floor(length(scaf.lines)/2)+(1:100)]
    png( filename=file.path(thisdir,"snp_map%03d.png"), 
            width=5*144, height=5*144, pointsize=10, res=144 )
    map_alleles( do.lines )
    dev.off()
    system(paste("cd", thisdir, ";",
                 file.path(tortdir,"visualization/make-slideshow.sh"),
                 "*.png > index.html" ) )
}
```
```{r map_outlier_scafs, include=FALSE, cache=TRUE}
for (scaf in outliers) {
    scafdir <- paste0(scaf,"_maps")
    dir.create(scafdir,showWarnings=FALSE)
    thisdir <- file.path( scafdir, "high_cor" )
    dir.create(thisdir,showWarnings=FALSE)
    scaf.lines <- which( pos$chr==scaf )
    do.lines <- scaf.lines[ ( order(pc1.df$snp.cor[scaf.lines],decreasing=TRUE) <= 100 ) & (pos$totDepth[scaf.lines] > mindepth) ]
    png( filename=file.path(thisdir,"snp_map%03d.png"), 
            width=5*144, height=5*144, pointsize=10, res=144 )
    map_alleles( do.lines )
    dev.off()
    system(paste("cd", thisdir, ";",
                 file.path(tortdir,"visualization/make-slideshow.sh"),
                 "*.png > index.html" ) )
}
```
Long scaffolds:
```{r map_scafs_list, results="asis", echo=FALSE}
list_to_mdlist <- function (x) { cat('\n',paste0('- ',paste(x,collapse='\n- ')),'\n',sep='','\n') }
list_to_mdlist( paste0('[', superlong, '](', superlong, '_maps/centered/index.html', ')') )
```

Scaffolds with many high-correlation SNPs:
```{r map_scafs_list_outliers, results="asis", echo=FALSE}
list_to_mdlist( paste0('[', outliers, '](', outliers, '_maps/high_cor/index.html', ')') )
```


# Maps of high-correlation SNPs

## With PC1

We'll make maps of all the SNPs that have correlation with PC1 above 0.85 
(there are `r sum(abs(pc1.df$snp.cor[( pos$totDepth >= mindepth ) & ( pos$totDepth <= maxdepth ) ])>0.85)` of these).
```{r high_cor_maps, include=FALSE, cache=TRUE, depends="snp_info"}
pc.num <- 1
high.cor.lines <- which( ( pos$totDepth >= mindepth ) & ( pos$totDepth <= maxdepth )  & abs(pc1.df$snp.cor)>0.85 )
thisdir <- sprintf("cor_pc%d_gt_0.85_maps",pc.num)
dir.create(thisdir,showWarnings=FALSE)
png( filename=file.path(thisdir,"snp_map%03d.png"), 
        width=5*144, height=5*144, pointsize=10, res=144 )
map_alleles( high.cor.lines )
dev.off()
system(paste("cd", thisdir, ";",
             file.path(tortdir,"visualization/make-slideshow.sh"),
             "*.png > index.html" ) )
```
These are in [`r thisdir`](`r file.path(thisdir,"index.html")`).

## With PC2

We'll also make maps of all the SNPs that have correlation with PC2 above 0.5 
(there are `r sum(abs(pc2.df[["snp.cor"]][( pos$totDepth >= mindepth ) & ( pos$totDepth <= maxdepth ) ])>0.5)` of these).
```{r high_cor_maps_2, include=FALSE, cache=TRUE, depends="snp_info"}
pc.num <- 2
high.cor.lines <- which( ( pos$totDepth >= mindepth ) & ( pos$totDepth <= maxdepth )  & abs(pc2.df$snp.cor)>0.5 )
thisdir <- sprintf("cor_pc%d_gt_0.5_maps",pc.num)
dir.create(thisdir,showWarnings=FALSE)
png( filename=file.path(thisdir,"snp_map%03d.png"), 
        width=5*144, height=5*144, pointsize=10, res=144 )
map_alleles( high.cor.lines )
dev.off()
system(paste("cd", thisdir, ";",
             file.path(tortdir,"visualization/make-slideshow.sh"),
             "*.png > index.html" ) )
```
These are in [`r thisdir`](`r file.path(thisdir,"index.html")`).


# Unfiltered?

There are a lot of SNPs that *don't* pass our filters that show very high correlation with PC1.
Many of these may have been removed only for being on shortish contigs 
(we required total coverage between `r mindepth` and `r maxdepth`, 
and at least `r minscaf` SNPs per contig).
Here is a plot of correlation with PC1 against total depth,
```{r all_snps, cache=TRUE, depends="snp_info"}
usethese <- ( !is.na(abs(pc1.df$snp.cor)) & abs(pc1.df$snp.cor)>0.7 )
plot( pos$totDepth[usethese], abs(pc1.df$snp.cor[usethese]), pch=20, cex=0.5, log='x', xlim=c(10,550),
     col=adjustcolor("black",0.5),
    xlab='total depth', ylab='correlation with PC1' )
```
Looks like most of the high correlation ones are due to *low* coverage,
i.e., just noise.


# Haplotypes

Now let's get some length statistics on possible haplotypes.
First we restrict to only scaffolds having correlation with PC1 at least 0.8.
```{r which_highcor}
high.cor.snps <- which( ( pos$totDepth >= mindepth ) & ( pos$totDepth <= maxdepth ) & (abs(pc1.df$snp.cor)>0.8) )
high.cor.scafs <- unique( pos$chr[high.cor.snps] )
```
There are `r length(high.cor.scafs)` of these, 
with total length `r sum(scaf.len[match(high.cor.scafs,levels(pos$chr))])`,
with is `r sum(scaf.len[match(high.cor.scafs,levels(pos$chr))])/sum(scaf.len)` of the genome.
Together they have `r nrow(subset(pos,chr%in%high.cor.scafs))` SNPs,
a density of `r nrow(subset(pos,chr%in%high.cor.scafs))/sum(scaf.len[match(high.cor.scafs,levels(pos$chr))])*1000` per Kb.
For comparison, the 16.5Kb of mitochondria has 34 SNPs with correlation above 0.8,
a density of 2 per Kb.
Here are all their SNPs:
```{r highcor_scafs}
scafplot( ifelse(seq_len(nrow(pos))%in%high.cor.snps,abs(pc1.df$snp.cor),NA), scaf=high.cor.scafs )
```
