---
title: "SNP clinality"
date: "`r date()`"
---

```{r setup, include=FALSE}
library(pander)
library(raster)
library(maptools)
# library(phonTools)  # for fastacf
options(scipen=3)
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
# locations and labels
tortdir <- gsub("tortoisescape.*","tortoisescape",getwd())
coord.obj <- load(file.path(tortdir,"tort_272_info","geog_coords.RData"))
coords <- get(coord.obj)
tort.ids <- row.names(coords)
tort.names <- gsub(" .*","",tort.ids)  # remove the " (sheared)"
nindivs <- length(tort.ids)
```

```{r snp_info, include=FALSE, cache=TRUE}
datadir <- "."
bincountfile <- file.path(datadir,"272torts_snp1e6_minmapq20minq30_map2kenro.counts.bin")
persitefile <- file.path(datadir,"272torts_snp1e6_minmapq20minq30_map2kenro.counts.bin.persite.stats")
perindivfile <- file.path(datadir,"272torts_snp1e6_minmapq20minq30_map2kenro.counts.perindividual.hists")
posfile <- file.path(datadir,"272torts_snp1e6_minmapq20minq30_map2kenro.pos.gz")
# maf <- read.table(file.path(datadir,"272torts_snp1e6_minmapq20minq30_map2kenro.mafs.gz"),header=TRUE)
persite <- read.table(persitefile,header=TRUE)
perindiv <- read.table(perindivfile,header=TRUE)
colnames(perindiv) <- tort.names
pos <- read.table(posfile,header=TRUE)
# scaffolds
scaf.counts <- table( pos$chr )
```

```{r count_connection, include=FALSE}
# the counts themselves
bincount <- file(bincountfile,open="rb")
attr(bincount,"nindivs") <- nindivs
attr(bincount,"nbytes") <- 1
read_bincounts <- function (bincount,lines) {
    line.coords <- 4*(lines-1)*attr(bincount,"nbytes")*attr(bincount,"nindivs")
    output <- matrix( integer(4*length(lines)*attr(bincount,"nindivs")), nrow=length(lines) )
    for (k in seq_along(lines)) {
        seek(bincount,line.coords[k])
        output[k,] <- readBin( bincount, what=integer(),
                              n=4*attr(bincount,"nindivs"),
                              size=attr(bincount,"nbytes"),
                              signed=(attr(bincount,"nbytes")>2) )
    }
    output[output>=256^attr(bincount,"nbytes")-1] <- NA
    return(output)
}
```

```{r checks, include=FALSE}
stopifnot( nrow(persite)==nrow(pos) )
```

There are `r nrow(persite)` total SNPs.
We first compute the total coverage at all sites except those in the largest depth bin
(greater than per-individual coverage `r nrow(perindiv)-1`):
```{r calcs}
mean.coverage <- colSums( sweep(perindiv,1,c(0:(nrow(perindiv)-2),0),"*") )/nrow(persite)
```

# Per-site statistics

First we should filter out bad-looking sites.
Here are distributions of the statistics we have:
```{r persite_hists}
nInd.table <- table(persite$nInd)
max.coverage.table <- table(persite$max_coverage)
min.coverage.table <- table(persite$min_coverage)
total.coverage <- persite$total_A + persite$total_C + persite$total_G + persite$total_T
total.coverage.table <- table(total.coverage)

plot( nInd.table, main="number of individuals covered",
     xlab='number of individuals', ylab='number of sites' )
plot( as.numeric(names(max.coverage.table)), as.numeric(max.coverage.table), log='xy',
     xlab='max coverage', ylab='number of sites' )
plot( as.numeric(names(min.coverage.table)), as.numeric(min.coverage.table), log='xy',
     xlab='min coverage', ylab='number of sites' )
plot( as.numeric(names(total.coverage.table)), as.numeric(total.coverage.table), log='xy',
     xlab='total coverage', ylab='number of sites' )
plot( as.numeric(names(total.coverage.table)), as.numeric(total.coverage.table), xlim=c(0,800),
     xlab='total coverage', ylab='number of sites' )
hist(log(persite$sd_coverage),breaks=100, main='SD of per-individual coverage')
hist(persite$heterozygosity,breaks=100, main='mean heterozygosity')

do.these <- ( total.coverage < 600 & persite$heterozygosity > 0 )
plot( hexbin::hexbin( total.coverage[do.these], persite$sd_coverage[do.these], xbins=100 ) )
plot( hexbin::hexbin( total.coverage[do.these], persite$heterozygosity[do.these], xbins=100 ) )
plot( hexbin::hexbin( persite$sd_coverage[do.these], persite$heterozygosity[do.these], xbins=100 ) )

```



# Are coverages Poisson?

Short answer: **no.**

Here are the distributions of coverage by individual:
```{r perindiv}
matplot( 0:(nrow(perindiv)-1), perindiv, type='l', log='y',
       xlab='coverage', ylab='number of sites' )
matplot( 0:(nrow(perindiv)-1), perindiv, type='l', xlim=c(0,30),
       xlab='coverage', ylab='number of sites' )
```
For each, we'll fit a Poisson distribution:
using Poisson probabilities, we expect
$p_k = e^{-\lambda} \lambda^k/k!$
and so $\log (k! p_k) = -\lambda + k \log \lambda$;
in other words, if coverage were Poisson, these plots would be linear:
```{r pos_fit}
cbins <- 0:40
matplot( cbins, (log(perindiv[1+cbins,])+lfactorial(cbins)), type='l',
        xlab='coverage', ylab='log(k! * number of sites)' )
cbins <- 2:10
matplot( cbins, (log(perindiv[1+cbins,])+lfactorial(cbins)), type='l',
        xlab='coverage', ylab='log(k! * number of sites)' )
```
Here's estimated $\lambda$ from various segments of bins:
```{r est_pos}
maxk.list <- 3:30
est.covs <- do.call( rbind, lapply( maxk.list, function (maxk) {
        cbins <- 2:maxk
        coverage.lms <- apply( log(perindiv[1+cbins,])+lfactorial(cbins), 2, function (x) coef(lm( x ~ cbins )) )
        return(exp(coverage.lms[2,]))
    } ) )
matplot( maxk.list, est.covs, xlab='coverage bin cutoff', ylab='estimated lambda', type='l' )
```
And here are those estimates from bins 2-3 and 2-4 against total coverage per individual:
```{r est_pos_vs_cov}
layout(t(1:2))
plot(mean.coverage, est.covs[1,])
abline(0,1)
plot(mean.coverage, est.covs[2,])
abline(0,1)
```
This doesn't seem to be getting anywhere.
- 

## Test of the method

Here's a quick check the method above works with simulated data.


```{r test_pois}
true.lambdas <- sort(4*runif(20))
nobs <- 1e6
maxcount <- 40
x <- do.call( cbind, lapply( true.lambdas, function (z) { tabulate(1+rpois(nobs,z),nbins=maxcount+1) } ) )
p <- sweep(x,2,colSums(x),"/")
exp.p <- outer(0:maxcount, true.lambdas, function (a,b) { exp(-b)*(b^a)/factorial(a) })
plot(as.vector(p),exp.p)
abline(0,1)
plot(log(p)+lfactorial(0:maxcount), outer(0:maxcount, true.lambdas, function (a,b) { -b + a*log(b) } ) )
abline(0,1)
```


For each, we'll fit a Poisson distribution:
using Poisson probabilities, we expect
$p_k = e^{-\lambda} \lambda^k/k!$
and so $\log (k! p_k) = -\lambda + k \log \lambda$;
in other words, if coverage were Poisson, these plots would be linear:
```{r pos_fit}


cbins <- 0:40
matplot( cbins, (log(x[1+cbins,])+lfactorial(cbins)), type='l',
        xlab='coverage', ylab='log(k! * number of sites)' )
matplot( cbins, outer( cbins, true.lambdas, function (a,b) { -b + a*log(b) } ), type='l' )

cbins <- 2:10
matplot( cbins, (log(x[1+cbins,])+lfactorial(cbins)), type='l',
        xlab='coverage', ylab='log(k! * number of sites)' )
matplot( cbins, outer( cbins, true.lambdas, function (a,b) { -b + a*log(b) } ), type='l' )

```
This looks pretty good for 2-5, with a different regime clearly taking over past then.
Here's total coverage per individual against the estimated $\lambda$ from these bins:
```{r est_pos}
maxk.list <- 3:10
est.covs <- do.call( rbind, lapply( maxk.list, function (maxk) {
        cbins <- 2:maxk
        coverage.lms <- apply( log(x[1+cbins,])+lfactorial(cbins), 2, function (z) coef(lm( z[is.finite(z)] ~ cbins[is.finite(z)] )) )
        return(exp(coverage.lms[2,]))
    } ) )
matplot( maxk.list, est.covs, xlab='coverage bin cutoff', ylab='estimated lambda', type='l' )


cbins <- 2:5
coverage.lms <- apply( log(x[1+cbins,])+lfactorial(cbins), 2, function (x) coef(lm( x ~ cbins )) )
plot( true.lambdas, exp(coverage.lms[2,]), 
     xlab='mean total coverage', ylab='poisson mean estimate' )
abline(0,1)
```
- 

