```{r setup, include=FALSE}
source("../resistance-fns.R")
require(parallel)
require(colorspace)
require(raster)
fig.dim <- 5
opts_chunk$set(fig.width=2*fig.dim,fig.align='center')
inflate <- function (...,fac=.2) { xr <- range(...,na.rm=TRUE); mean(xr) + (1+fac) * (xr-mean(xr)) }
# read config
config <- read.json.config("report-config.json")
layer.names <- config$layer_names
for (x in config$setup_files) { load( x ) }
# set up 
# plotting stuff
ph <- plot.ht.fn("../../geolayers/multigrid/256x/crm_",nonmissing,homedir="../..",default.par.args=list(mar=c(3,3,3,3)+.1))
layer <- with( environment(ph), layer )
sample.coords <- with(environment(ph),coordinates(tort.coords.rasterGCS))
```

Another method of interpolating hitting times
=============================================

First, compute the full matrix of true hitting times, which we assume are observed only on a well-spread-out subset of the space:
```{r true_hts, cache=TRUE}
ref.inds <- which.nonoverlapping(neighborhoods)
obs.locs <- locs[ref.inds]
true.params <- paramvec(config)
true.G <- G
true.G@x <- update.G(true.params)
true.hts <- hitting.analytic(neighborhoods[ref.inds],true.G)
#  and subset out the observed ones:
true.obs.hts <- true.hts[locs[ref.inds],]
```

We interpolate by letting $P$ be the projection matrix from the full hitting times to the observed ones, and finding $H$ to minimize
$$
\| GH + 1 \|^2 + \alpha \|PH - T\|^2
$$
where $\alpha$ is a tuning parameter and $T$ are the observed hitting times.
The minimizer of this solves
$$
\left( G^T G + \alpha P^T P \right) H = P^T T - G^T 1 
$$

Now, to interpolate these.
Here are the solutions using `solve()`, for $\alpha=0$ and $\alpha=1$:
```{r interp_hts, fig.width=2*fig.dim, fig.height=fig.dim, cache=TRUE}
interp.hts.0 <- interp.hitting( neighborhoods[ref.inds], true.G, true.obs.hts, obs.locs=locs[ref.inds], alpha=0 )
interp.hts.1 <- interp.hitting( neighborhoods[ref.inds], true.G, true.obs.hts, obs.locs=locs[ref.inds], alpha=1 )
layout(t(1:2))
plot(1+as.vector(true.hts), 1+as.vector(interp.hts.0), log='xy', xlab="true hitting times", ylab="interpolated hitting times", main=expression(alpha==0)); abline(0,1)
plot(1+as.vector(true.hts), 1+as.vector(interp.hts.1), log='xy', xlab="true hitting times", ylab="interpolated hitting times", main=expression(alpha==1)); abline(0,1)
```

Now, let's try Jacobi iteration.
Pulling from `interp.hitting`,
```{r setup_jacobi, fig.width=fig.dim, fig.height=fig.dim}
G <- true.G
jacobi.problem <- function (k,alpha) {
    klocs <- neighborhoods[[ref.inds[k]]]
    obs.ht <- true.obs.hts
    G.d <- G - Diagonal(nrow(G),rowSums(G))
    # Pmat projects full hitting times onto the obs.locs
    Pmat <- sparseMatrix( i=seq_along(obs.locs), j=obs.locs, x=1, dims=c(length(obs.locs),nrow(G)) )
    PtP <- alpha * crossprod(Pmat)
    bvec <- as.vector( alpha * crossprod(Pmat[,-klocs],obs.ht[,k]) + crossprod( G.d[-klocs,-klocs], rep(-1.0,nrow(G)-length(klocs)) ) )
    GP <- PtP[-klocs,-klocs] + crossprod(G[-klocs,-klocs])
    return( list( k=k, klocs=klocs, Dinv=1/diag(GP), R=GP-Diagonal(nrow(GP),diag(GP)), bvec=bvec ) )
}
jacobi.iter <- function (ht,jp) {
    as.vector( (jp$Dinv) * ( jp$bvec - jp$R %*% ht ) )
}
jp <- jacobi.problem(1,alpha=1)
# check this works
plot( true.hts[-jp$klocs,1], jacobi.iter(true.hts[-jp$klocs,1],jp) ); abline(0,1)
# and iterate
interp.mat <- matrix( 0, nrow=nrow(true.hts), ncol=20 )
interp.mat[-jp$klocs,1] <- mean(true.hts)
for (k in 2:ncol(interp.mat)) {
    interp.mat[-jp$klocs,k] <- jacobi.iter(interp.mat[-jp$klocs,k-1],jp)
}
matplot(t(sweep(interp.mat[,1:20],1,true.hts[,1],"-")),type='l')
```
No, that doesn't work.
Hmph.

Wait, let's try using some intuition here.
Instead, let's modify the Jacobi step that we'd get from solving $GH=-1$
to also keep the solution close to the observed hitting times:
define the matrix $R_{xy} = G_{xy}/(-G_{xx})$ and $R_{xx}=0$, which subtracts the local average;
so the Jacobi iteration to solve $GH=-1$ is:
replace $H_x$ by $1/(-G_{xx}) + RH_x$,
which is (time to jump from x) + (local average of H).
Instead, for sites where we have observed $T$, we can replace $H_x$ with
$(1-\alpha)(1/(-G_{xx}) + RH_x) + \alpha T_x$.
```{r setup_jacobi_2, fig.width=fig.dim, fig.height=fig.dim}
G <- true.G
jacobi.problem.2 <- function (k) {
    locs <- neighborhoods[ref.inds]
    obs.ht <- true.obs.hts[,k]
    dG <- rowSums(G)
    R <- G
    R@x <- G@x/dG[1L+G@i]
    klocs <- unlist(locs[k])[!is.na(unlist(locs[k]))]
    return( list( k=k, klocs=klocs, inv.dG=1/dG, R=R, obs.ht=obs.ht ) )
}
jacobi.iter.2 <- function (ht,jp,alpha) {
    z <- as.vector( 0.5 * ( ht + jp$inv.dG + jp$R%*%ht ) )
    z[ obs.locs ] <- (1-alpha)*z[ obs.locs ] + alpha*jp$obs.ht
    z[ jp$klocs ] <- 0
    return(z)
}
jp.2 <- jacobi.problem.2(1)
layout(t(1:2))
plot( 1+true.hts[,1], 1+jacobi.iter.2(true.hts[,1],jp.2,alpha=0), log='xy' ); abline(0,1)
plot( 1+true.hts[,1], 1+jacobi.iter.2(true.hts[,1],jp.2,alpha=0.1), log='xy' ); abline(0,1)
```
That works.  Iteration?

```{r iter_jacobi_2, fig.width=2*fig.dim, fig.height=fig.dim}
interp.mat <- matrix( 0, nrow=nrow(true.hts), ncol=500 )
interp.mat[,1] <- true.hts[,1] * exp( rnorm(nrow(true.hts),sd=1.0) )
for (k in 2:ncol(interp.mat)) {
    interp.mat[,k] <- jacobi.iter.2(interp.mat[,k-1],jp.2,alpha=1.0)
}
layout(t(1:2))
norm.error <- (interp.mat - true.hts[,1])/(1+true.hts[,1])
matplot( t( norm.error[unlist(neighborhoods),] ), type='l', ylim=range(norm.error[20,]) )

plot( 1+true.hts[,1], 1+interp.mat[,ncol(interp.mat)], log='xy', xlab='true hitting time', ylab='interpolated hitting time' )
points( (1+true.hts[,1])[unlist(neighborhoods)], (1+interp.mat[,ncol(interp.mat)])[unlist(neighborhoods)], col='red', pch=20 )
abline(0,1)
```

Here's what it's doing:
```{r jacobi_images, fig.width=fig.dim, fig.height=10*fig.dim}
layout(1:10)
for (k in floor(seq(1,ncol(interp.mat),length.out=10)) ) {
    # ph( log( (1+interp.mat[,k])/(1+true.hts[,1]) ) )
    ph( interp.mat[,k], zlim=inflate(range(true.hts[obs.locs,])), main=k)
}
```

```{r}
# layout(t(1:2))
# for (k in floor(seq(1,ncol(interp.mat),length.out=50)) ) {
#     # ph( log( (1+interp.mat[,k])/(1+true.hts[,1]) ) )
#     ph( interp.mat[,k], zlim=inflate(range(true.hts[obs.locs,])), main=k)
#     ph( true.hts[,1]-interp.mat[,k], zlim=(range(true.hts[obs.locs,]-interp.mat[obs.locs,50])), main=k)
#     if (length(locator(1))==0) { break; }
# }
```
