```{r setup}
source("../resistance-fns.R")
require(parallel)
require(colorspace)
require(raster)
fig.dim <- 5
opts_chunk$set(fig.width=2*fig.dim,fig.align='center')
show( load("cache/model-description-saved-state.RData") )   # saved in model-description.R
for (x in config$setup_files) { load(x) }
```

This continues from where [model-description.Rmd](model-description.html) leaves off.


Sensitivity of hitting times to the parameters
==============================================

Now, we only observe hitting times between the observed locations (points on the plots).
We can compute how much each parameter affects those hitting times.
First we compute the derivative of each hitting time plotted above with respect to each parameter,
which we can then visualize.
Here, rows are parameters; columns are hitting times to individuals shown in the previous plot;
white is zero, blue colors are negative, and red colors are positive.
```{r sensitivity, fig.width=4*fig.dim, fig.height=4*fig.dim, cache=TRUE}
hs <- hitting.sensitivity(params, neighborhoods[plot.inds], G, update.G, layers, transfn, valfn, ndelta, ngamma)
obs.hs <- sapply( hs, function (x) { x[locs,] } )
layout(matrix(1:(length(params)*(length(plot.inds)+1)),nrow=length(params),byrow=TRUE))
par(mar=c(0,0,0,0)+.1)
for (j in seq_along(params)) {
    plot(c(0,2), c(0,2), type='n', xaxt='n', yaxt='n')
    text(1,1,label=names(params)[j],offset=0)
    for (k in seq_along(plot.inds)) {
        ph( hs[[j]][,k], xaxt='n', yaxt='n', par.args=list(mar=c(0,0,0,0)+.1), legend=FALSE, 
            zlim=1.2*c(-1,1)*max(abs(obs.hs)), col=diverge_hcl(255) )
    }
}
```
We see the hitting times are least sensitive to the `delta` parameters,
and that hitting times to the second individual (who is isolated) are the most sensitive.

Which direction can we change the parameters in while changing the observed hitting times the least?
First, let's compute the covariance matrix of the parameters:
```{r param_covariance, fig.width=fig.dim, fig.height=fig.dim}
hs <- hitting.sensitivity(params, neighborhoods[ref.inds], G, update.G, layers, transfn, valfn, ndelta, ngamma)
param.cov <- hitting.colinearity(obs.locs=locs[ref.inds],hs=hs)
image(Matrix(param.cov))
# direction of minimal and maximal change:
min.dirn <- eigen(param.cov)$vectors[,length(params)]
max.dirn <- eigen(param.cov)$vectors[,1]
names(min.dirn) <- names(max.dirn) <- names(params)
rbind( min=min.dirn, max=max.dirn )
```
Now let's see how perturbing the parameters in these directions changes the hitting times:
```{r min_max_paramchange, cache=TRUE, fig.height=fig.dim}
d.hts <- function (params) {
    G@x <- update.G(params)
    mean( (hitting.analytic(neighborhoods[ref.inds],G)[locs[ref.inds],] - hts[locs[ref.inds],ref.inds])^2 )
}
epsvals <- 0.1*seq(-1,1,length.out=11)
min.d.hts <- sapply( epsvals, function (eps) { d.hts( params + eps * min.dirn ) } )
max.d.hts <- sapply( epsvals, function (eps) { d.hts( params + eps * max.dirn ) } )
layout(t(1:2))
plot( epsvals, min.d.hts, xlab="perturbation", ylab="mean hitting time difference", main='minimal direction' )
plot( epsvals, max.d.hts, xlab="perturbation", ylab="mean hitting time difference", main='maximal direction' )
```
Note the change in scale: changing in the maximal direction changes the hitting times about $10^3$ more quickly than in the minimal direction.



Sensitivity of best-fit parameters to the hitting times
=======================================================

Noiseless hitting times
-----------------------

Suppose we know, without error,
pairwise hitting times between the reference sampling locations.
Do these uniquely determine the parameters?
Below, we vary each parameter, with the others fixed at their true values,
and plot the mean squared difference between the hitting times at the new parameter value
and at the original, "true" value.
Here is the plot for $\beta$:
```{r parameter_basin, cache=TRUE, fig.height=fig.dim, fig.width=fig.dim}
c.beta <- compute.plot.nearby( d.hts, params, fac=.1, do.params=1, npoints=11 )
plot.nearby( computed=c.beta )
```
and the plots for $\gamma$:
```{r gamma_basin, cache=TRUE, fig.height=2*fig.dim}
c.gamma <- compute.plot.nearby( d.hts, params, fac=.1, do.params=1+(1:6), npoints=11 )
layout( matrix(1:6,ncol=2) )
plot.nearby( computed=c.gamma )
```
and for $\delta$:
```{r delta_basin, cache=TRUE, fig.height=2*fig.dim}
c.delta <- compute.plot.nearby( d.hts, params, fac=.1, do.params=7+(1:6), npoints=11 )
layout( matrix(1:6,ncol=2) )
plot.nearby( computed=c.delta )
```

Noisy hitting times
-------------------

This is very promising; it means that we should have good power to infer the parameters --
under the model, the true parameters clearly give the best fit to the observed hitting times.
Now, what if we add a little noise to the hitting times?
Here, let's add varying amounts of noise,
and look at the same plots for three of the parameters.
```{r fuzz_basin, cache=TRUE, fig.height=fig.dim}
epsvals <- c(.01,.1)
fuzz.nearby <- lapply( epsvals, function (eps) {
        hts <- hts * exp( rnorm(length(hts))*eps )
        environment(d.hts) <- environment()  # what happens in lapply stays in lapply
        compute.plot.nearby( d.hts, params, fac=.2, do.params=c(1,2,8), npoints=11 )
    } )
```
First, with hitting times fuzzed by 1%, the best fit parameters hardly change:
```{r fuzz_basin_1, cache=TRUE, fig.height=fig.dim}
cat("fuzzed by ", epsvals[1], "\n")
layout(t(1:3))
plot.nearby( computed=fuzz.nearby[[1]] )
```
With the parameters fuzzed more dramatically, by 10%, they change somewhat more:
```{r fuzz_basin_2, cache=TRUE, fig.height=fig.dim}
cat("fuzzed by ", epsvals[2], "\n")
layout(t(1:3))
plot.nearby( computed=fuzz.nearby[[2]] )
```
but it is still overall quite promising.
