```{r setup, include=FALSE}
source("../resistance-fns.R")
require(parallel)
require(colorspace)
require(raster)
fig.dim <- 5
opts_chunk$set(fig.width=2*fig.dim,fig.align='center')
inflate <- function (...,fac=.2) { xr <- range(...,na.rm=TRUE); mean(xr) + (1+fac) * (xr-mean(xr)) }
# read config
config <- read.json.config("report-config.json")
layer.names <- config$layer_names
for (x in config$setup_files) { load( x ) }
# set up 
# plotting stuff
ph <- plot.ht.fn("../../geolayers/multigrid/256x/crm_",nonmissing,homedir="../..",default.par.args=list(mar=c(3,3,3,3)+.1))
layer <- with( environment(ph), layer )
sample.coords <- with(environment(ph),coordinates(tort.coords.rasterGCS))
# more setup from inference-procedure.Rmd
ref.inds <- which.nonoverlapping(neighborhoods)
true.params <- paramvec(config)
true.G <- G
true.G@x <- update.G(true.params)
true.hts <- hitting.analytic(neighborhoods[ref.inds],true.G)
#  and subset out the observed ones:
true.obs.hts <- true.hts[locs[ref.inds],]
```



Hitting time interpolation
==========================

The first thing that often happens when we try to interpolate hitting times
using a randomly chosen set of parameters is a numerical error (the matrix is nearly singular).
It turns out (below) that switching from the matrix $G$ to its symmetrized version helps,
perhaps because this lets us `solve()` using a Cholesky decomposition rather than the LU decomposition.

Singular matrices?
------------------

Here is the error:
```{r singular_LU}
init.params <- c( 1.9, -1.4, -1.0, 1.6, 0.4, -2.1, 2.8, -2.0, -0.2, 0.5, -0.5, -2.1, -0.9 )  # these make a singular matrix
G@x <- update.G(init.params)
check.inds <- c(9,19)
# can find hitting times in the usual way
wrong.hts <- hitting.analytic( neighborhoods[ref.inds[check.inds]], G )
# but not the associated quadratic interpolation problem
try( { wrong.interp.hts <- interp.hitting( neighborhoods[ref.inds[check.inds]], G, true.obs.hts[check.inds,check.inds], obs.locs=locs[ref.inds[check.inds]], alpha=0, numcores=1 ) } )
invisible(G[-1,1])
```
Since $\alpha=0$, the two problems should be the same.

Let's try with the symmetrized version of $G$, i.e.\ $J_{ij} = G_{ij} \sqrt{\pi_j/\pi_i}$. 
Let $Z$ be the diagonal matrix with $\sqrt{\pi}$ on the diagonal,
so that $J=Z^{-1} G Z$.
Then if $y = Z^{-1}x$, and $\nu_i = 1/\sqrt{\pi_i}$, then
$$
  G x = ZJy = Z \nu = -1  \iff Jy = -\nu .
$$
Note that $J_{ii} = G_{ii}$.
```{r try_symm, fig.width=fig.dim, fig.height=fig.dim}
pivec <- stationary.dist( init.params, layers, transfn )
# try with the symmetrized matrix
J <- G
J@x <- G@x * ( sqrt(pivec)[Gjj] / sqrt(pivec)[1L+G@i] )
range(J-t(J)) # check this is actually numerically symmetric
J <- forceSymmetric(J)
nu <- 1/sqrt(pivec)
# check this will give us hitting times
sym.hts <- numeric(nrow(G))
kn <- neighborhoods[[ref.inds[1]]]
sym.hts[-kn] <- as.vector( solve( (J-Diagonal(nrow(G),rowSums(G)))[-kn,-kn], (-1)*nu[-kn] ) * sqrt(pivec[-kn]) )
orig.hts <- hitting.analytic( neighborhoods[ref.inds[1]], G )
range( ((G-Diagonal(nrow(G),rowSums(G)))%*%sym.hts)[-kn] )
range( ((G-Diagonal(nrow(G),rowSums(G)))%*%orig.hts)[-kn] )
plot( 1+orig.hts, 1+sym.hts, log='xy'); abline(0,1)
```

That checks out.  Now, let's plug $J$ in for $G$ in the interpolation routine:
```{r interp_sym, fig.width=fig.dim, fig.height=fig.dim}
sym.bvec <- as.vector( crossprod( (J-Diagonal(nrow(G),rowSums(G)))[-kn,-kn], (-1)*nu[-kn] ) )
sym.hts.2 <- numeric(nrow(G))
sym.hts.2[-kn] <-  sqrt(pivec[-kn]) * as.vector( solve( crossprod( (J-Diagonal(nrow(G),rowSums(G)))[-kn,-kn] ), sym.bvec ) )
range( ((G-Diagonal(nrow(G),rowSums(G)))%*%sym.hts.2)[-kn] )
plot( 1+orig.hts, 1+sym.hts.2, log='xy'); abline(0,1)
```
That looks pretty good, although the solution is somewhat less good than before.

Does the coded up function agree?
```{r}
test.sym <- interp.hitting.sym(neighborhoods[ref.inds[1]], J, pivec, true.obs.hts[,1,drop=FALSE], obs.locs, alpha=0)
range( (test.sym-sym.hts.2)/test.sym, finite=TRUE )
```
Hm, to within 0.003, not that close.


Now, to interpolate, we want to find $x$ so that $Gx=-1$ and $x[i]=t[i]$;
or if $y=Z^{-1}x$, 
we want $Jy=-\nu$ and $y[i]=Zt[i]$.
```{r interp_sym_1, fig.width=fig.dim, fig.height=fig.dim}
test.interp <- interp.hitting.sym(neighborhoods[ref.inds[1]], J, pivec, true.obs.hts[,1,drop=FALSE], obs.locs, alpha=1e8)
plot(1+true.hts[,k], 1+test.interp, log='xy', pch=20, col=adjustcolor("black",.2), xlab="true hitting times", ylab="interpolated hitting times", main=expression(alpha==1)); abline(0,1)
points( 1+as.vector(true.obs.hts[,k]), 1+as.vector(test.interp[obs.locs]), col='red', pch=20 )
```

OK, works.
