```{r setup, include=FALSE}
if (!exists("infile")) { stop("Need to specify an input file.") }
source("resistance-fns.R")
require(parallel)
require(colorspace)
require(raster)
load(infile)
fig.dim <- 5
opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
inflate <- function (...,fac=.2) { xr <- range(...,na.rm=TRUE); mean(xr) + (1+fac) * (xr-mean(xr)) }
config <- read.json.config(trust.optim$config.file)
layer.names <- c("constant",config$layer_names)
prev.file <- if (is.null(trust.optim$prev.file)) { "None" } else { trust.optim$prev.file}
prev.html <- gsub(".RData",".html",basename(prev.file))
nlayers <- (length(trust.optim$argument)-2)/2
```

The inference procedure produced by
```
`r paste(trust.optim$invocation,collapse=' ')`
```
produced
- continuing from [`r prev.file`](`r prev.html`)
- iterations: `r trust.optim$iterations`
- converged: `r trust.optim$converged`
- acceptance rate: `r mean(trust.optim$accept)`
- mean step size: `r mean(trust.optim$stepnorm)`  


The fitting procedure:
```{r descent, fig.height=.75*fig.dim, echo=FALSE}
plot( trust.optim$valpath, col=2-trust.optim$accept, ylab="value", main="value", xlab="step number", type='b', log='y' )
plot( trust.optim$stepnorm, ylab="step size", main="step size", xlab="step number", type='b' )
plot( trust.optim$r, type='b', xlab='step number', ylab='radius', main='radius' )
plot( trust.optim$rho, xlab='step number', ylab='rho', main='ratio of predicted to observed improvement', type='b', ylim=c(-2,2) )
```

And here is how the parameters went:
```{r params}
matplot( trust.optim$argpath[,1,drop=FALSE], type='l', ylab='shift parameter', xlab='step number' )
matplot( trust.optim$argpath[,-1], type='l', ylab='parameter values', xlab='step number', lty=c(1,rep(1,nlayers),rep(2,nlayers)), col=c(1,2:(1+nlayers),2:(1+nlayers)) )
legend("topleft", lty=1, col=1:(1+nlayers), legend=c('beta',layer.names))
```

Now, let's look at the actual model.
```{r model_setup, echo=FALSE}
for (x in config$setup_files) { load(file.path(dirname(trust.optim$config.file),x)) }
ph <- plot.ht.fn(nonmissing=nonmissing,layer=nalayer,homedir="..")
params <- trust.optim$argument
layout(t(1:2))
plot.model(params,layer.names,layers,G,update.G,ph)
```

Now, hitting times:
```{r hitting_times, fig.width=3*fig.dim, fig.height=3*fig.dim, echo=FALSE}
G@x <- update.G(params[-1])
hts <- hitting.analytic(neighborhoods,G)
plot.inds <- c(1,2,16,23,27,54,77,98,107)
layout(matrix(1:9,nrow=3))
par(mar=c(0,0,2,0)+.1)
for (k in plot.inds) {
    ph( hts[,k], main=k, xaxt='n', yaxt='n', xlab='', ylab='' )
}
```

Here are the fitted hitting times against the values we're trying to fit,
and how the fit compares to distance:
```{r fit_hitting_times, echo=FALSE}
sym.hts <- params[1]+hts[locs,]
sym.hts <- (sym.hts+t(sym.hts))/2
ref.inds <- trust.optim$ref.inds
diag(sym.hts) <- diag(pimat) <- NA
tort.dists <- pointDistance(sample.locs,lonlat=FALSE)
ut <- upper.tri(pimat,diag=FALSE)
layout(t(1:2))
plot( sym.hts[ut], pimat[ut], col=adjustcolor(ifelse(col(pimat)[ut]%in%ref.inds,'red','black'),0.25), pch=20, cex=0.5, ylab="observed times", xlab="fitted times" )
legend("bottomright",pch=20,col=c("red","black"),legend=c("used in fitting","not used in fitting"))
abline(0,1)
plot( tort.dists[ut], pimat[ut], col=adjustcolor(ifelse(col(pimat)[ut]%in%ref.inds,'red','black'),0.25), pch=20, cex=0.5, ylab="observed times", xlab="geographic distance" )
```

